---
title: "Available methods in selectInferToolkit Package"
author: "Jinal Shah & Ryan A Peterson"
date: "`r Sys.Date()`"
output:   
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Available methods in selectInferToolkit Package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE)
```

# Overview

The `selectInferToolkit` package provides a unified interface for **post-selection inference** (PSI), statistical inference that accounts for model selection. Ignoring the model selection while performing inference leads to confidence intervals that are too narrow and p-values that are anti-conservative, even when the selection and inference are performed on the same dataset.

This vignette is designed to document which combinations of selection method, inference
method, and tuning options are currently supported by the package. Rather than maintaining a static table by hand, the vignette probes every combination
automatically each time it is rendered: it fits a small synthetic dataset, attempts every pipeline, and records whether each combination succeeds, produces a warning, or fails.  The compatibility matrix and flow diagrams below therefore
always aim to reflect the current state of the package.

## Selection methods

The package currently supports four families of variable/model selection, each
configurable with tuning parameters:

| Function | Method | Key options |
|---|---|---|
| `select_full_model()` | No selection and all variables retained | — |
| `select_stepwise_ic()` | Forward / backward / bi-directional stepwise | `penalty` (AIC / BIC), `direction` |
| `select_glmnet()` | Lasso / elastic net / ridge via `glmnet` | `alpha` (0 = ridge, between 0 and 1 = elastic net, 1 = lasso), `lambda` |
| `select_ncvreg()` | MCP or lasso via `ncvreg` | `penalty` (MCP / lasso), `lambda` |

## Inference methods

Given a selected model object, the package offers four inference approaches:

| Function | Method | Key options |
|---|---|---|
| `infer_upsi()` | Unajusted post-selection inference (UPSI)  | `nonselection` |
| `infer_selective()` | Selective inference: exact conditioning on the selection event; currently requires lasso or forward stepwise selection | `nonselection` |
| `infer_boot()` | Bootstrap-based inference: resamples the full selection + estimation pipeline | `inference_target`, `debias`, `estimation_data` |
| `infer_pipe()` | | — |

## The `nonselection` argument

For `infer_upsi()` and `infer_selective()`, variables that were *not* selected
present a choice about how to handle their inference:

- **`"ignored"`** - only report CIs for selected variables; unselected variables are dropped from the output entirely.
- **`"confident_nulls"`** - treat unselected variables as having point estiamtes and CI as exactly zero; report a point mass at zero with no uncertainty.
- **`"uncertain_nulls"`** - aims to make inference on all variables but without the assumption that non-selections certainly have a null effect


```{r packages}
library(selectInferToolkit)
library(DiagrammeR)
library(DT)
library(dplyr)
library(tidyr)
library(purrr)
library(hdrm)
```




```{r candidates}
####### Candidate space #######
# ── Auto-discover any new select_* exports ────────────────────────────────────
pkg_ns  <- tryCatch(getNamespace("selectInferToolkit"), error = function(e) NULL)
base_sel <- c("select_full_model","select_stepwise_ic","select_glmnet","select_ncvreg")
if (!is.null(pkg_ns)) {
  base_sel <- union(base_sel, grep("^select_", ls(pkg_ns), value=TRUE))
}

# ── Selection option grids ────────────────────────────────────────────────────
# One row = one "flavour" (function + specific tuning values).
# sel_label always prefixes the function name so rows are unambiguous across
# selectors, e.g. "select_glmnet | alpha=1" vs "select_ncvreg | penalty=lasso".
sel_grids <- list(

  select_full_model = data.frame(sel_params="", stringsAsFactors=FALSE),

  select_stepwise_ic = expand.grid(
    penalty   = c("AIC", "BIC"),
    direction = c("forward", "backward", "both"),
    stringsAsFactors = FALSE
  ),

  select_glmnet = expand.grid(
    alpha  = c(0, 0.5, 1),   # 0=ridge, 0.5=elastic net, 1=lasso
    lambda = c("best", "compact"),
    stringsAsFactors = FALSE
  ),

  select_ncvreg = expand.grid(
    penalty = c("MCP", "lasso"),
    lambda  = c("best", "compact"),
    stringsAsFactors = FALSE
  )
)

sel_grids <- imap(sel_grids, function(g, nm) {
  param_cols <- setdiff(names(g), c("sel_label","sel_fn","sel_params"))
  if (length(param_cols) == 0 || (ncol(g)==1 && names(g)=="sel_params")) {
    g$sel_label  <- nm          # no tuning: label = function name only
    g$sel_params <- ""
  } else {
    g$sel_params <- apply(g[, param_cols, drop=FALSE], 1,
                          function(r) paste(names(r), r, sep="=", collapse=" | "))
    g$sel_label  <- paste0(nm, " | ", g$sel_params)
  }
  g$sel_fn <- nm
  g
})

# ── Inference option spaces ───────────────────────────────────────────────────
nonselection_values <- c("ignored", "confident_nulls", "uncertain_nulls")
inf_fns_nonsel      <- c("infer_selective", "infer_upsi")

boot_option_grid <- expand.grid(
  inference_target = c("selections", "all"),
  debias           = c(TRUE, FALSE),
  estimation_data  = c("in-sample"),
  stringsAsFactors = FALSE
) %>%
  mutate(boot_label = paste0("target=", inference_target,
                             " | debias=", debias,
                             " | est=", estimation_data))

# ── Colour palettes ───────────────────────────────────────────────────────────
inf_palette <- c(infer_boot="#F18F01", infer_selective="#2E86AB", infer_upsi="#A23B72", infer_pipe="#2D6A4F")
sel_palette <- c(select_full_model="#546E7A", select_stepwise_ic="#6D4C41",
                 select_glmnet="#2E7D32",    select_ncvreg="#4527A0")
```




```{r probe, cache=TRUE}
####### Probe every combination #######


set.seed(42)
data <- hdrm::gen_data_abn(n=150, p=10, a=2, b = 2, rho = .8,
                             beta = c(0.3, -0.4))
x=data$X
y= data$y
df= data.frame(y=data$y, data$X)


# ── probe_status: run expr_fn(), return supported/supported_with_warning/not_supported
probe_status <- function(expr_fn) {
  warn_msg <- NULL
  result <- withCallingHandlers(
    tryCatch(
      { expr_fn(); list(status="supported", message=NA_character_) },
      error = function(e) list(status="not_supported", message=conditionMessage(e))
    ),
    warning = function(w) { warn_msg <<- conditionMessage(w); invokeRestart("muffleWarning") }
  )
  if (result$status == "supported" && !is.null(warn_msg))
    result <- list(status="supported_with_warning", message=warn_msg)
  result
}

run_sel <- function(sel_fn, extra_args=list()) {
  fn <- get(sel_fn, envir=getNamespace("selectInferToolkit"))
  do.call(fn, c(list(y ~., data=df), extra_args))
}

# Flat tibble: one row per selector flavour, extra_args stored as list-column.
# IMPORTANT: use split()+lapply() not apply() — apply() coerces the whole
# data frame to a character matrix first, turning numeric alpha=0.5 into
# the string "0.5". That lets select_glmnet silently accept alpha="0.5"
# instead of erroring, so infer_selective fails to flag the incompatibility.
all_sel_rows <- imap_dfr(sel_grids, function(g, sel_fn) {
  param_cols <- setdiff(names(g), c("sel_label","sel_fn","sel_params"))
  args_list  <- if (length(param_cols) == 0) {
    vector("list", nrow(g))
  } else {
    # split() into 1-row data frames then as.list() preserves column types
    lapply(split(g[, param_cols, drop=FALSE], seq_len(nrow(g))), as.list)
  }
  tibble(sel_fn=sel_fn, sel_label=g$sel_label, sel_params=g$sel_params,
         extra_args=args_list)
})

#  Grid 2a: infer_selective / infer_upsi × selector flavours × nonselection 
inf_nonsel_combos <- expand.grid(
  inference=inf_fns_nonsel, nonselection=nonselection_values,
  stringsAsFactors=FALSE
)
n_sel <- nrow(all_sel_rows); n_inf <- nrow(inf_nonsel_combos)

grid_nonsel <- all_sel_rows[rep(seq_len(n_sel), each=n_inf), ] %>%
  bind_cols(inf_nonsel_combos[rep(seq_len(n_inf), times=n_sel), ]) %>%
  mutate(boot_label=NA_character_)

probes_nonsel <- map(seq_len(nrow(grid_nonsel)), function(i) {
  probe_status(function() {
    sel <- run_sel(grid_nonsel$sel_fn[i], grid_nonsel$extra_args[[i]])
    fn  <- get(grid_nonsel$inference[i], envir=getNamespace("selectInferToolkit"))
    fn(sel, nonselection=grid_nonsel$nonselection[i], data=df)
  })
})

results_nonsel <- grid_nonsel %>%
  mutate(status  = map_chr(probes_nonsel, "status"),
         message = map_chr(probes_nonsel, ~.x$message %||% NA_character_))

#  Grid 2b: infer_boot × selector flavours × boot option grid 
n_boot <- nrow(boot_option_grid)

grid_boot <- all_sel_rows[rep(seq_len(n_sel), each=n_boot), ] %>%
  bind_cols(boot_option_grid[rep(seq_len(n_boot), times=n_sel), ]) %>%
  mutate(inference="infer_boot", nonselection=NA_character_)

probes_boot <- map(seq_len(nrow(grid_boot)), function(i) {
  probe_status(function() {
    sel <- run_sel(grid_boot$sel_fn[i], grid_boot$extra_args[[i]])
    infer_boot(sel, data=df,
               inference_target=grid_boot$inference_target[i],
               debias=grid_boot$debias[i],
               estimation_data=grid_boot$estimation_data[i],
               B=5)
  })
})

results_boot <- grid_boot %>%
  mutate(status  = map_chr(probes_boot, "status"),
         message = map_chr(probes_boot, ~.x$message %||% NA_character_))

# Grid 2c: infer_pipe × selector flavours (no nonselection, no option grid)
# infer_pipe takes only the model object and data — no extra arguments.
grid_pipe <- all_sel_rows %>%
  mutate(inference="infer_pipe", nonselection=NA_character_, boot_label=NA_character_)

probes_pipe <- map(seq_len(nrow(grid_pipe)), function(i) {
  probe_status(function() {
    sel <- run_sel(grid_pipe$sel_fn[i], grid_pipe$extra_args[[i]])
    infer_pipe(sel, data=df)
  })
})

results_pipe <- grid_pipe %>%
  mutate(status  = map_chr(probes_pipe, "status"),
         message = map_chr(probes_pipe, ~.x$message %||% NA_character_))

# Combine all three probe grids
results <- bind_rows(results_nonsel, results_boot, results_pipe) %>%
  select(sel_fn, sel_params, sel_label, inference, nonselection, boot_label,
         status, message) %>%
  arrange(sel_fn, inference, sel_params, nonselection, boot_label)
```




# Compatibility matrix

The table below summarises which pipelines are currently functional. Each row is one *selector flavor* - a specific model selection function with a
particular combination of tuning parameters (e.g. `select_glmnet` with `alpha=1.0, lambda=best`).  Each column group is one inference method.

**How to read the table:**

- For `infer_selective` and `infer_upsi`, the cell shows the consensus result
  across all three `nonselection` values.  If all three behave the same way the cell shows a single icon; if they differ a tally is shown (e.g. ✅×2 ❌×1).
- For `infer_boot`, results are split by `estimation_data` (in-sample vs
  out-of-sample) but collapsed across `inference_target` × `debias`, since
  currently those options do not affect whether the method runs successfully.
- For `infer_pipe`, there are no tuning options - the cell simply indicates
  whether the method accepts output from that selector.
  


**Notable constraints visible in the table:**

- `infer_selective` only works with lasso-penalised glmnet selection
  (`alpha = 1.0`) and forward stepwsie selection.  
- `infer_pipe` works with `select_ncvreg` but not with stepwise or glmnet
  selectors.
- `infer_upsi` and `infer_boot` are the most broadly compatible, working across all selector families.


```{r matrix-combined}
icon_of <- function(s) case_when(
  s=="supported"              ~ "\u2705",
  s=="supported_with_warning" ~ "\u26a0\ufe0f",
  TRUE                        ~ "\u274c"
)

# Collapse helper: if all icons agree return that icon, else tally
consensus_icon <- function(icons) {
  icons <- icons[!is.na(icons)]
  if (length(icons) == 0) return("\u274c")
  tbl <- sort(table(icons), decreasing=TRUE)
  if (length(tbl) == 1) return(names(tbl))
  paste(paste0(names(tbl), "\u00d7", as.integer(tbl)), collapse=" ")
}

# ── Part 1: infer_selective & infer_upsi ─────────────────────────────────────
# Collapse all 3 nonselection values into ONE column per inference method.
# Almost always the same result for all three, shown as a single icon.
# A footnote explains that all three nonselection values were tested.
mat_nonsel <- results %>%
  filter(inference %in% c("infer_selective", "infer_upsi")) %>%
  mutate(icon = icon_of(status)) %>%
  group_by(sel_fn, sel_params, inference) %>%
  summarise(cell = consensus_icon(icon), .groups="drop") %>%
  pivot_wider(names_from=inference, values_from=cell)

# ── Part 2: infer_boot collapsed to 2 columns (in-sample / out-of-sample) ────
# Each cell = consensus across all 4 inference_target × debias combos.
mat_boot_wide <- results %>%
  filter(inference == "infer_boot") %>%
  mutate(
    est_data = ifelse(grepl("out-of-sample", boot_label, fixed=TRUE),
                      "infer_boot\n(out-of-sample)", "infer_boot\n(in-sample)"),
    icon = icon_of(status)
  ) %>%
  group_by(sel_fn, sel_params, est_data) %>%
  summarise(cell = consensus_icon(icon), .groups="drop") %>%
  pivot_wider(names_from=est_data, values_from=cell)

# ── Part 3: infer_pipe — single column, no options ───────────────────────────
mat_pipe <- results %>%
  filter(inference == "infer_pipe") %>%
  mutate(icon = icon_of(status)) %>%
  group_by(sel_fn, sel_params) %>%
  summarise(infer_pipe = consensus_icon(icon), .groups="drop")

# ── Join all parts ────────────────────────────────────────────────────────────
mat_combined <- mat_nonsel %>%
  left_join(mat_boot_wide, by=c("sel_fn","sel_params")) %>%
  left_join(mat_pipe,      by=c("sel_fn","sel_params")) %>%
  rename(`selector fn`=sel_fn, `selector params`=sel_params)

datatable(
  mat_combined,
  rownames=FALSE, escape=FALSE,
  caption=paste(
    "\u2705 supported | \u26a0\ufe0f warning | \u274c not supported. |"
  ),
  options=list(
    pageLength=30, dom="t", scrollX=TRUE,
    columnDefs=list(list(className="dt-center", targets="_all"))
  )
)
```


Note: `infer_upsi` and `infer_selective` columns show consensus across all three `nonselection` values (`ignored`, `confident_nulls`, `uncertain_nulls`).`infer_boot` columns show consensus across all `inference_target` and `debias` combinations, split by estimation_data.

## Flow diagrams

The diagrams below visualise the same information as the compatibility matrix
but as a directed graph, making it easier to trace which full pipelines are
available for a given selector.

**How to read the diagrams:**

- **Left column** - selector flavours (specific function + tuning parameter
  combination).  All nodes in this column share the same colour, indicating
  they belong to the same selector family.
- **Middle column** - inference methods.  Each method has its own colour,
  carried through to its edges and terminal nodes.
- **Right column** - supported option values for the chosen inference method
  (nonselection values for `infer_upsi`/`infer_selective`; option dimensions
  for `infer_boot`; "(no options)" for `infer_pipe`).
- **Solid edges** - the combination is fully supported.
- **Missing edges** - the combination is not supported and is excluded from
  the diagram.

One diagram is shown per selector family.  `select_full_model` is omitted.


```{r diagram-helpers, include=FALSE}
# ── status_rank: used to summarise across option variants ─────────────────────
# Purpose: for each (selector flavour × inference method) pair there may be
# multiple option variants (e.g. 3 nonselection values). We want to know the
# *best* outcome across all those variants so we can decide whether to draw
# an edge at all (any support = draw edge) and whether it should be solid or
# dashed (solid if any variant is fully supported, dashed if only warnings).
# Combinations where every variant fails get rank 1 and are excluded from the
# diagram, but they remain visible in the full results table in §5.
status_rank <- c(supported=3L, supported_with_warning=2L, not_supported=1L)

pair_best <- results %>%
  mutate(rank = status_rank[status]) %>%
  group_by(sel_fn, sel_label, inference) %>%
  summarise(best_rank=max(rank), .groups="drop") %>%
  mutate(best_status = names(status_rank)[match(best_rank, status_rank)]) %>%
  filter(best_rank >= 2L)   # keep only pairs where ≥1 variant works

# ── Compute which specific option values are supported per (sel_fn × inference)
# For nonselection methods: list supported nonselection values
# For infer_boot: list supported combinations of inference_target, debias, est_data
supported_nonsel_vals <- results %>%
  filter(!inference %in% c("infer_boot","infer_pipe"), status %in% c("supported","supported_with_warning")) %>%
  group_by(sel_fn, inference) %>%
  summarise(vals = list(sort(unique(nonselection))), .groups="drop")
# (kept for reference; make_diagram now queries `results` directly per value)


supported_boot_vals <- results %>%
  filter(inference == "infer_boot", status %in% c("supported","supported_with_warning")) %>%
  separate_wider_delim(
    boot_label,
    delim = " | ",
    names = c("target", "debias", "est")
  ) %>%
  mutate(
    inference_target = sub("target=", "", target),
    debias = sub("debias=", "", debias),
    estimation_data  = sub("est=", "", est)
  ) %>%
  group_by(sel_fn) %>%
  summarise(
    targets  = list(sort(unique(inference_target))),
    debiases = list(sort(unique(as.character(debias)))),
    est_data = list(sort(unique(estimation_data))),
    .groups  = "drop"
  )

# ── DOT helpers ───────────────────────────────────────────────────────────────
safe_id <- function(x) gsub("[^a-zA-Z0-9_]", "_", x)

node_def <- function(id, label, fill, fontcolor="white",
                     fontsize=10, width=1.9, height=0.45, penwidth=0,
                     border_col=fill) {
  sprintf(
    '  "%s" [label="%s", shape="rectangle", style="filled,rounded",
             fillcolor="%s", fontcolor="%s", fontsize=%d,
             width=%.2f, height=%.2f, penwidth=%.1f, color="%s"]',
    safe_id(id), gsub('"', "'", label), fill, fontcolor,
    fontsize, width, height, penwidth, border_col
  )
}

edge_def <- function(from, to, col, dashed=FALSE, penwidth=1.5) {
  sprintf('  "%s" -> "%s" [color="%s", style="%s", penwidth=%.1f, arrowsize=0.6]',
          safe_id(from), safe_id(to), col,
          if (dashed) "dashed" else "solid", penwidth)
}

# ── make_diagram ───────────────────────────────────────────────────────────────
# Selector flavour nodes are coloured by their selector family (sel_palette).
# Inference method nodes and edges use inf_palette — one colour per method.
# Terminal option nodes use a faded version of their inference method colour.

make_diagram <- function(sel_fn_filter, pair_best, inf_palette, sel_palette,
                         results, supported_boot_vals,
                         splines = "polyline") {
  # splines="ortho" crashes Graphviz maze router (maze.c assertion) on dense
  # graphs. "polyline" is stable at any size. Pass splines="curved" for a
  # slightly prettier look on smaller diagrams if desired.

  pb <- pair_best %>% filter(sel_fn == sel_fn_filter)
  if (nrow(pb) == 0) return(invisible(NULL))

  sel_flavours <- unique(pb$sel_label)
  inf_shown    <- unique(pb$inference)
  sel_col      <- sel_palette[[sel_fn_filter]]

  lines <- c(
    'digraph {',
    sprintf('graph [layout="dot", rankdir="LR", splines="%s",', splines),
    '       bgcolor="white", nodesep=0.35, ranksep=0.8]',
    'node [fontname="Helvetica"]',
    'edge [fontname="Helvetica", fontsize=8]',
    '',
    '// Column 1: selector flavour nodes (each its own colour)'
  )

  for (s in sel_flavours) {
    disp <- sub(paste0("^", sel_fn_filter, "\\s*\\|\\s*"), "", s)
    if (disp == s) disp <- s
    lines <- c(lines, node_def(s, disp, fill=sel_col, fontcolor="white",
                               width=2.2, penwidth=0))
  }

  # Inference nodes: coloured by inf_palette
  lines <- c(lines, '', '// Column 2: inference method nodes (inf_palette colour)')
  for (i in inf_shown) {
    lines <- c(lines, node_def(i, i, fill=inf_palette[[i]], fontcolor="white",
                               width=1.9, penwidth=0))
  }

  # Terminal nodes: one per nonselection value OR one per boot option dimension.
  # These are shared across flavours — the colour story is told by the edges.
  lines <- c(lines, '', '// Column 3: terminal option value nodes (neutral)')

  for (i in inf_shown) {
    # Terminal nodes use a faded version of the inference method's own colour
    term_fill <- paste0(inf_palette[[i]], "BB")

    if (i == "infer_boot") {
      bv <- supported_boot_vals %>% filter(sel_fn == sel_fn_filter)
      if (nrow(bv) == 0) next
      dims <- list(
        list(id    = paste0(safe_id(i), "_target"),
             label = paste0("inference_target\n",
                            paste(bv$targets[[1]], collapse=" | "))),
        list(id    = paste0(safe_id(i), "_debias"),
             label = paste0("debias\n",
                            paste(bv$debiases[[1]], collapse=" | "))),
        list(id    = paste0(safe_id(i), "_estdata"),
             label = paste0("estimation_data\n",
                            paste(bv$est_data[[1]], collapse=" | ")))
      )
      for (d in dims) {
        lines <- c(lines, sprintf(
          '  "%s" [label="%s", shape="rectangle", style="filled,rounded",
                   fillcolor="%s", fontcolor="white", fontsize=8,
                   width=2.1, height=0.55, penwidth=0]',
          d$id, d$label, term_fill))
      }

    } else if (i == "infer_pipe") {
      # infer_pipe has no option arguments — single terminal node
      lines <- c(lines, sprintf(
        '  "%s" [label="(no options)", shape="rectangle", style="filled,rounded",
                 fillcolor="%s", fontcolor="white", fontsize=8,
                 width=1.6, height=0.42, penwidth=0]',
        paste0(safe_id(i), "_noopts"), term_fill))

    } else {
      # nonselection methods (infer_selective, infer_upsi)
      supported_rows <- results %>%
        filter(sel_fn==sel_fn_filter, inference==i,
               status %in% c("supported","supported_with_warning"))
      nonsel_vals_used <- sort(unique(supported_rows$nonselection))
      if (length(nonsel_vals_used) == 0) next
      for (v in nonsel_vals_used) {
        nid <- paste0(safe_id(i), "_", safe_id(v))
        lines <- c(lines, sprintf(
          '  "%s" [label="%s", shape="rectangle", style="filled,rounded",
                   fillcolor="%s", fontcolor="white", fontsize=8,
                   width=1.9, height=0.42, penwidth=0]',
          nid, v, term_fill))
      }
    }
  }

  # ── Edges: flavour → inference (coloured by inference method) ────────────────
  lines <- c(lines, '', '// Flavour → inference edges (inf_palette colour)')
  for (s in sel_flavours) {
    infs_ok <- pb %>% filter(sel_label == s) %>% pull(inference) %>% unique()
    for (i in infs_ok) {
      best <- pb %>% filter(sel_label==s, inference==i) %>% pull(best_status)
      dashed <- best == "supported_with_warning"
      lines <- c(lines, edge_def(s, i, inf_palette[[i]], dashed, penwidth=1.6))
    }
  }

  # ── Edges: inference → terminal (coloured by inference method)
  # Drawn once per inference method (not per flavour) since colour is per-method.
  lines <- c(lines, '', '// Inference → terminal edges (inf_palette colour)')
  for (i in inf_shown) {
    col <- inf_palette[[i]]

    if (i == "infer_boot") {
      bv <- supported_boot_vals %>% filter(sel_fn == sel_fn_filter)
      if (nrow(bv) == 0) next
      for (sfx in c("_target","_debias","_estdata")) {
        lines <- c(lines, edge_def(i, paste0(safe_id(i), sfx), col, penwidth=1.0))
      }

    } else if (i == "infer_pipe") {
      lines <- c(lines, edge_def(i, paste0(safe_id(i), "_noopts"), col, penwidth=1.0))

    } else {
      # nonselection methods: one edge per supported nonselection value
      supported_rows <- results %>%
        filter(sel_fn==sel_fn_filter, inference==i,
               status %in% c("supported","supported_with_warning"))
      for (v in sort(unique(supported_rows$nonselection))) {
        lines <- c(lines, edge_def(i, paste0(safe_id(i),"_",safe_id(v)), col, penwidth=1.0))
      }
    }
  }

  lines <- c(lines, '}')
  grViz(paste(lines, collapse="\n"))
}
```



### `select_stepwise_ic`  (penalty × direction)

```{r diag-stepwise, fig.width=11, fig.height=6}
make_diagram("select_stepwise_ic", pair_best, inf_palette, sel_palette,
             results, supported_boot_vals, splines = "ortho")
```


**Interpretation:** All six AIC/BIC × direction combinations support
`infer_upsi` and `infer_boot`.  `infer_selective` is only available for
`direction = "forward"` — this reflects the fact that the selective inference
conditioning set is only tractable when the forward-stepwise path is used.
`infer_pipe` is not available for stepwise selection.

### `select_glmnet`  (alpha × lambda: ridge/elastic net/lasso × best/compact)

```{r diag-glmnet, fig.width=11, fig.height=5}
make_diagram("select_glmnet", pair_best, inf_palette, sel_palette,
             results, supported_boot_vals,splines = "ortho")
```


**Interpretation:** `infer_selective` (shown in blue) is only connected for
`alpha = 1.0` (lasso) flavours.  Ridge (`alpha = 0`) and elastic net
(`alpha = 0.5`) can only be paired with `infer_upsi` or `infer_boot`.
All `lambda` options work equivalently for supported inference methods.


### `select_ncvreg`  (penalty × lambda)

```{r diag-ncvreg, fig.width=11, fig.height=5}
make_diagram("select_ncvreg", pair_best, inf_palette, sel_palette,
             results, supported_boot_vals)
```

**Interpretation:** Both MCP and lasso penalties, and both `lambda` options,
support `infer_upsi`and `infer_boot`.  `infer_selective` and `infer_pipe` is
not available for `ncvreg` selectors.


## For developers

This vignette is designed to update itself automatically as new methods are
added to the package.  The probing logic in the `probe` chunk calls each
function via `get(fn_name, envir=getNamespace("selectInferToolkit"))`, so
any newly exported `select_*` or `infer_*` function is discovered without
editing this file.

To add a **new selection method** with tuning parameters, add one entry to
`sel_grids` in the `candidates` chunk:

```r
select_mymethod = expand.grid(
  option1 = c("a", "b"),
  option2 = c(TRUE, FALSE),
  stringsAsFactors = FALSE
)
```

To add a **new inference method that uses `nonselection=`**, add its name to
`inf_fns_nonsel`.

To add a **new inference method with its own option grid** (like `infer_boot`), add a dedicated probe block in the `probe` chunk following the pattern used for `grid_boot`.

The compatibility matrix and flow diagrams will include the new method
automatically on the next render.



```{r full-table, include=FALSE}

## Full results table
datatable(
  results %>%
    rename(`selector fn`=sel_fn, `selector params`=sel_params,
           `selector label`=sel_label, `boot options`=boot_label),
  rownames=FALSE, filter="top",
  options=list(pageLength=15, scrollX=TRUE),
  caption="All probed combinations — filter by any column"
)
```






