---
title: "Available methods in selectInferToolkit Package"
author: "Jinal Shah & Ryan Peterson"
date: "`r Sys.Date()`"
output:   
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Available methods in selectInferToolkit Package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE)
```

# Overview

The `selectInferToolkit` package provides a unified interface for **post-selection inference** (PSI), statistical inference that accounts for model selection. Ignoring the model selection while performing inference leads to confidence intervals that are too narrow and p-values that are anti-conservative, as the selection and inference are performed on the same dataset.

This vignette is designed to document which combinations of selection method, inference
method, and tuning options are currently supported by the package. Rather than maintaining a static table by hand, the vignette probes every combination
automatically each time it is rendered: it fits a small synthetic dataset, attempts every pipeline, and records whether each combination succeeds, produces a warning, or fails.  The compatibility matrix and flow diagrams below therefore
reflect the state of the package at the time this vignette was built.

The probe dataset uses the "A / B / N" simulation design of Patrick Breheny,
implemented as `gen_data_abn()` in the
[hdrm](https://github.com/pbreheny/hdrm) package: `n = 150` observations on
`p = 10` predictors, arranged as two blocks of three exchangeably-correlated
variables (`rho = 0.8`) plus four independent noise variables.  Within each
block the `A` variable has a nonzero coefficient (0.3 and -0.4) while the two
`B` variables are correlated with it but have coefficients of exactly zero —
precisely the correlated-but-null predictors that selection procedures tend to
pick up by mistake.  The design is reproduced in base R inside this vignette
rather than taken from `hdrm` only so that the package carries no non-CRAN
dependencies.

## Selection methods

The package currently supports four families of variable/model selection, each
configurable with tuning parameters:

| Function | Method | Key options |
|---|---|---|
| `select_full_model()` | No selection and all variables retained | — |
| `select_stepwise_ic()` | Forward / backward / bi-directional stepwise | `penalty` (AIC / BIC), `direction` |
| `select_glmnet()` | Lasso / elastic net / ridge via `glmnet` | `alpha` (0 = ridge, between 0 and 1 = elastic net, 1 = lasso), `lambda` |
| `select_ncvreg()` | MCP, SCAD, or lasso via `ncvreg` | `penalty` (MCP / SCAD / lasso), `lambda` |

## Inference methods

Given a selected model object, the package offers four inference approaches:

| Function | Method | Key options |
|---|---|---|
| `infer_upsi()` | Unadjusted post-selection inference (UPSI) | `nonselection` |
| `infer_selective()` | Selective inference: exact conditioning on the selection event; currently requires lasso or forward stepwise selection | `nonselection` |
| `infer_boot()` | Bootstrap-based inference: resamples the full selection + estimation pipeline | `inference_target`, `debias`, `estimation_data` |
| `infer_pipe()` | PIPE: projection-based test statistics and intervals for a penalized fit (wraps `ncvreg::intervals()`); currently requires `ncvreg` selection, and is experimental | — |

## The `nonselection` argument

For `infer_upsi()` and `infer_selective()`, variables that were *not* selected
present a choice about how to handle their inference:

- **`"ignored"`** - only selected variables receive inference; non-selected variables are still listed in the output, but their estimates and CIs are returned as `NA`.
- **`"confident_nulls"`** - treat non-selected variables as exact zeros: point estimate and both CI limits are zero, i.e. a point mass at zero with no uncertainty.
- **`"uncertain_nulls"`** - aims to make inference on all variables, conditioning on the selected model and checking each non-selection against residuals of that model individually.


```{r packages}
library(selectInferToolkit)
library(DiagrammeR)
library(DT)
library(dplyr)
library(tidyr)
library(purrr)
```




```{r candidates}
####### Candidate space #######
# ── Auto-discover any new select_* exports ────────────────────────────────────
pkg_ns  <- tryCatch(getNamespace("selectInferToolkit"), error = function(e) NULL)
base_sel <- c("select_full_model","select_stepwise_ic","select_glmnet","select_ncvreg")
if (!is.null(pkg_ns)) {
  base_sel <- union(base_sel, grep("^select_", ls(pkg_ns), value=TRUE))
}

# ── Selection option grids ────────────────────────────────────────────────────
# One row = one "flavor" (function + specific tuning values).
# sel_label always prefixes the function name so rows are unambiguous across
# selectors, e.g. "select_glmnet | alpha=1" vs "select_ncvreg | penalty=lasso".
sel_grids <- list(

  select_full_model = data.frame(sel_params="", stringsAsFactors=FALSE),

  select_stepwise_ic = expand.grid(
    penalty   = c("AIC", "BIC"),
    direction = c("forward", "backward", "both"),
    stringsAsFactors = FALSE
  ),

  select_glmnet = expand.grid(
    alpha  = c(0, 0.5, 1),   # 0=ridge, 0.5=elastic net, 1=lasso
    lambda = c("best", "compact"),
    stringsAsFactors = FALSE
  ),

  select_ncvreg = expand.grid(
    penalty = c("MCP", "SCAD", "lasso"),
    lambda  = c("best", "compact"),
    stringsAsFactors = FALSE
  )
)

sel_grids <- imap(sel_grids, function(g, nm) {
  param_cols <- setdiff(names(g), c("sel_label","sel_fn","sel_params"))
  if (length(param_cols) == 0 || (ncol(g)==1 && names(g)=="sel_params")) {
    g$sel_label  <- nm          # no tuning: label = function name only
    g$sel_params <- ""
  } else {
    g$sel_params <- apply(g[, param_cols, drop=FALSE], 1,
                          function(r) paste(names(r), r, sep="=", collapse=" | "))
    g$sel_label  <- paste0(nm, " | ", g$sel_params)
  }
  g$sel_fn <- nm
  g
})

# ── Inference option spaces ───────────────────────────────────────────────────
nonselection_values <- c("ignored", "confident_nulls", "uncertain_nulls")
inf_fns_nonsel      <- c("infer_selective", "infer_upsi")

boot_option_grid <- expand.grid(
  inference_target = c("selections", "all"),
  debias           = c(TRUE, FALSE),
  estimation_data  = c("in-sample", "out-of-sample"),
  stringsAsFactors = FALSE
) %>%
  mutate(boot_label = paste0("target=", inference_target,
                             " | debias=", debias,
                             " | est=", estimation_data))

# ── Color palettes ───────────────────────────────────────────────────────────
inf_palette <- c(infer_boot="#F18F01", infer_selective="#2E86AB", infer_upsi="#A23B72", infer_pipe="#2D6A4F")
sel_palette <- c(select_full_model="#546E7A", select_stepwise_ic="#6D4C41",
                 select_glmnet="#2E7D32",    select_ncvreg="#4527A0")
```




```{r probe}
####### Probe every combination #######


set.seed(42)

# Synthetic "A / B / N" design, due to Patrick Breheny and implemented as
# hdrm::gen_data_abn() in the hdrm package (https://github.com/pbreheny/hdrm).
# It is reproduced here in base R only so that this vignette carries no
# non-CRAN dependencies; the design itself is Breheny's.
#
# Two blocks of three exchangeably-correlated predictors: within each block the
# "A" variable carries a nonzero coefficient and the two "B" variables are
# correlated with it but have coefficients of exactly zero.  Four further "N"
# variables are independent noise.  The B variables are what make this a useful
# probe -- they are the correlated-but-null predictors that selection methods
# most often pick up by mistake.
n <- 150; a <- 2; b <- 2; rho <- 0.8; n_noise <- 4
beta_a <- c(0.3, -0.4)

# Exchangeable block: all pairwise correlations equal rho, all variances 1.
gen_block <- function(n, k, rho) {
  shared <- rnorm(n)
  sqrt(rho) * shared + sqrt(1 - rho) * matrix(rnorm(n * k), n, k)
}

X <- cbind(
  do.call(cbind, lapply(seq_len(a), function(i) gen_block(n, b + 1, rho))),
  matrix(rnorm(n * n_noise), n, n_noise)
)
colnames(X) <- c(
  unlist(lapply(seq_len(a),
                function(i) c(paste0("A", i), paste0("B", (i - 1) * b + seq_len(b))))),
  paste0("N", seq_len(n_noise))
)

beta <- setNames(numeric(ncol(X)), colnames(X))
beta[paste0("A", seq_len(a))] <- beta_a
y  <- as.vector(X %*% beta) + rnorm(n)
df <- data.frame(y = y, X)


# ── probe_status: run expr_fn(), return supported/supported_with_warning/not_supported
probe_status <- function(expr_fn) {
  warn_msg <- NULL
  result <- withCallingHandlers(
    tryCatch(
      { expr_fn(); list(status="supported", message=NA_character_) },
      error = function(e) list(status="not_supported", message=conditionMessage(e))
    ),
    warning = function(w) { warn_msg <<- conditionMessage(w); invokeRestart("muffleWarning") }
  )
  if (result$status == "supported" && !is.null(warn_msg))
    result <- list(status="supported_with_warning", message=warn_msg)
  result
}

run_sel <- function(sel_fn, extra_args=list()) {
  # Re-seed before every fit. cv.glmnet() and cv.ncvreg() draw fresh CV folds
  # each time they are called, so the *same* selector flavor can select
  # different variables in different probes. That makes cells flicker between
  # renders and produces spurious tallies (e.g. "✅×2 ⚠️×1") that reflect CV
  # noise rather than any real difference between the options being compared.
  set.seed(2024)
  fn <- get(sel_fn, envir=getNamespace("selectInferToolkit"))
  do.call(fn, c(list(y ~., data=df), extra_args))
}

# Flat tibble: one row per selector flavor, extra_args stored as list-column.
# IMPORTANT: use split()+lapply() not apply() — apply() coerces the whole
# data frame to a character matrix first, turning numeric alpha=0.5 into
# the string "0.5". That lets select_glmnet silently accept alpha="0.5"
# instead of erroring, so infer_selective fails to flag the incompatibility.
all_sel_rows <- imap_dfr(sel_grids, function(g, sel_fn) {
  param_cols <- setdiff(names(g), c("sel_label","sel_fn","sel_params"))
  args_list  <- if (length(param_cols) == 0) {
    vector("list", nrow(g))
  } else {
    # split() into 1-row data frames then as.list() preserves column types
    lapply(split(g[, param_cols, drop=FALSE], seq_len(nrow(g))), as.list)
  }
  tibble(sel_fn=sel_fn, sel_label=g$sel_label, sel_params=g$sel_params,
         extra_args=args_list)
})

#  Grid 2a: infer_selective / infer_upsi × selector flavors × nonselection 
inf_nonsel_combos <- expand.grid(
  inference=inf_fns_nonsel, nonselection=nonselection_values,
  stringsAsFactors=FALSE
)
n_sel <- nrow(all_sel_rows); n_inf <- nrow(inf_nonsel_combos)

grid_nonsel <- all_sel_rows[rep(seq_len(n_sel), each=n_inf), ] %>%
  bind_cols(inf_nonsel_combos[rep(seq_len(n_inf), times=n_sel), ]) %>%
  mutate(boot_label=NA_character_)

probes_nonsel <- map(seq_len(nrow(grid_nonsel)), function(i) {
  probe_status(function() {
    sel <- run_sel(grid_nonsel$sel_fn[i], grid_nonsel$extra_args[[i]])
    fn  <- get(grid_nonsel$inference[i], envir=getNamespace("selectInferToolkit"))
    fn(sel, nonselection=grid_nonsel$nonselection[i], data=df)
  })
})

results_nonsel <- grid_nonsel %>%
  mutate(status  = map_chr(probes_nonsel, "status"),
         message = map_chr(probes_nonsel, ~.x$message %||% NA_character_))

#  Grid 2b: infer_boot × selector flavors × boot option grid 
n_boot <- nrow(boot_option_grid)

grid_boot <- all_sel_rows[rep(seq_len(n_sel), each=n_boot), ] %>%
  bind_cols(boot_option_grid[rep(seq_len(n_boot), times=n_sel), ]) %>%
  mutate(inference="infer_boot", nonselection=NA_character_)

probes_boot <- map(seq_len(nrow(grid_boot)), function(i) {
  probe_status(function() {
    sel <- run_sel(grid_boot$sel_fn[i], grid_boot$extra_args[[i]])
    infer_boot(sel, data=df,
               inference_target=grid_boot$inference_target[i],
               debias=grid_boot$debias[i],
               estimation_data=grid_boot$estimation_data[i],
               B=5)
  })
})

results_boot <- grid_boot %>%
  mutate(status  = map_chr(probes_boot, "status"),
         message = map_chr(probes_boot, ~.x$message %||% NA_character_))

# Grid 2c: infer_pipe × selector flavors (no nonselection, no option grid)
# infer_pipe takes only the model object and data — no extra arguments.
grid_pipe <- all_sel_rows %>%
  mutate(inference="infer_pipe", nonselection=NA_character_, boot_label=NA_character_)

probes_pipe <- map(seq_len(nrow(grid_pipe)), function(i) {
  probe_status(function() {
    sel <- run_sel(grid_pipe$sel_fn[i], grid_pipe$extra_args[[i]])
    infer_pipe(sel, data=df)
  })
})

results_pipe <- grid_pipe %>%
  mutate(status  = map_chr(probes_pipe, "status"),
         message = map_chr(probes_pipe, ~.x$message %||% NA_character_))

# Combine all three probe grids
results <- bind_rows(results_nonsel, results_boot, results_pipe) %>%
  select(sel_fn, sel_params, sel_label, inference, nonselection, boot_label,
         status, message) %>%
  arrange(sel_fn, inference, sel_params, nonselection, boot_label)
```




# Compatibility matrix

The table below summarizes which pipelines are currently functional. Each row is one *selector flavor* - a specific model selection function with a
particular combination of tuning parameters (e.g. `select_glmnet` with `alpha=1.0, lambda=best`).  Each column group is one inference method.

**How to read the table:**

- For `infer_selective` and `infer_upsi`, the cell shows the consensus result
  across all three `nonselection` values.  If all three behave the same way the cell shows a single icon; if they differ a tally is shown (e.g. ✅×2 ❌×1).
- For `infer_boot`, results are split into two columns by `estimation_data`
  (in-sample vs out-of-sample) but collapsed across `inference_target` ×
  `debias`, since currently those options do not affect whether the method
  runs successfully.
- For `infer_pipe`, there are no tuning options - the cell simply indicates
  whether the method accepts output from that selector.
  


**Notable constraints visible in the table:**

- `infer_selective` only works with lasso-penalized glmnet selection
  (`alpha = 1.0`) and forward stepwise selection.
- `infer_pipe` works with `select_ncvreg` but not with stepwise or glmnet
  selectors.
- `infer_upsi` and `infer_boot` are the most broadly compatible, working across all selector families.
- The out-of-sample `infer_boot` column is marked ⚠️ throughout: the method
  runs for every selector, but warns that out-of-sample estimation is
  experimental and needs a large sample size.  Treat those cells as "runs, but
  use with care" rather than "recommended".


```{r matrix-combined}
icon_of <- function(s) case_when(
  s=="supported"              ~ "\u2705",
  s=="supported_with_warning" ~ "\u26a0\ufe0f",
  TRUE                        ~ "\u274c"
)

# Collapse helper: if all icons agree return that icon, else tally
consensus_icon <- function(icons) {
  icons <- icons[!is.na(icons)]
  if (length(icons) == 0) return("\u274c")
  tbl <- sort(table(icons), decreasing=TRUE)
  if (length(tbl) == 1) return(names(tbl))
  paste(paste0(names(tbl), "\u00d7", as.integer(tbl)), collapse=" ")
}

# ── Part 1: infer_selective & infer_upsi ─────────────────────────────────────
# Collapse all 3 nonselection values into ONE column per inference method.
# Almost always the same result for all three, shown as a single icon.
# A footnote explains that all three nonselection values were tested.
mat_nonsel <- results %>%
  filter(inference %in% c("infer_selective", "infer_upsi")) %>%
  mutate(icon = icon_of(status)) %>%
  group_by(sel_fn, sel_params, inference) %>%
  summarize(cell = consensus_icon(icon), .groups="drop") %>%
  pivot_wider(names_from=inference, values_from=cell)

# ── Part 2: infer_boot collapsed to 2 columns (in-sample / out-of-sample) ────
# Each cell = consensus across all 4 inference_target × debias combos.
mat_boot_wide <- results %>%
  filter(inference == "infer_boot") %>%
  mutate(
    est_data = ifelse(grepl("out-of-sample", boot_label, fixed=TRUE),
                      "infer_boot\n(out-of-sample)", "infer_boot\n(in-sample)"),
    icon = icon_of(status)
  ) %>%
  group_by(sel_fn, sel_params, est_data) %>%
  summarize(cell = consensus_icon(icon), .groups="drop") %>%
  pivot_wider(names_from=est_data, values_from=cell)

# ── Part 3: infer_pipe — single column, no options ───────────────────────────
mat_pipe <- results %>%
  filter(inference == "infer_pipe") %>%
  mutate(icon = icon_of(status)) %>%
  group_by(sel_fn, sel_params) %>%
  summarize(infer_pipe = consensus_icon(icon), .groups="drop")

# ── Join all parts ────────────────────────────────────────────────────────────
mat_combined <- mat_nonsel %>%
  left_join(mat_boot_wide, by=c("sel_fn","sel_params")) %>%
  left_join(mat_pipe,      by=c("sel_fn","sel_params")) %>%
  rename(`selector fn`=sel_fn, `selector params`=sel_params)

datatable(
  mat_combined,
  rownames=FALSE, escape=FALSE,
  caption="\u2705 supported | \u26a0\ufe0f supported, with a warning | \u274c not supported",
  options=list(
    pageLength=30, dom="t", scrollX=TRUE,
    columnDefs=list(list(className="dt-center", targets="_all"))
  )
)
```


## Flow diagrams

The diagrams below visualize the same information as the compatibility matrix
but as a directed graph, making it easier to trace which full pipelines are
available for a given selector.

**How to read the diagrams:**

- **Left column** - selector flavors (specific function + tuning parameter
  combination).  All nodes in this column share the same color, indicating
  they belong to the same selector family.
- **Middle column** - inference methods.  Each method has its own color,
  carried through to its edges and terminal nodes.
- **Right column** - supported option values for the chosen inference method
  (nonselection values for `infer_upsi`/`infer_selective`; option dimensions
  for `infer_boot`; "(no options)" for `infer_pipe`).
- **Solid edges** - the combination is fully supported (at least one option
  variant runs without any warning).
- **Dashed edges** - the combination runs, but every option variant that works
  emits a warning; see the corresponding ⚠️ cell in the matrix above.
- **Missing edges** - the combination is not supported and is excluded from
  the diagram.

One diagram is shown per selector family.  `select_full_model` is omitted
because it has no tuning parameters, so its diagram would reduce to the single
row already shown in the matrix above.


```{r diagram-helpers, include=FALSE}
# ── status_rank: used to summarize across option variants ─────────────────────
# Purpose: for each (selector flavor × inference method) pair there may be
# multiple option variants (e.g. 3 nonselection values). We want to know the
# *best* outcome across all those variants so we can decide whether to draw
# an edge at all (any support = draw edge) and whether it should be solid or
# dashed (solid if any variant is fully supported, dashed if only warnings).
# Combinations where every variant fails get rank 1 and are excluded from the
# diagram, but they remain visible in the "All probed combinations" table at
# the end of this vignette.
status_rank <- c(supported=3L, supported_with_warning=2L, not_supported=1L)

pair_best <- results %>%
  mutate(rank = status_rank[status]) %>%
  group_by(sel_fn, sel_label, inference) %>%
  summarize(best_rank=max(rank), .groups="drop") %>%
  mutate(best_status = names(status_rank)[match(best_rank, status_rank)]) %>%
  filter(best_rank >= 2L)   # keep only pairs where ≥1 variant works

# ── Compute which specific option values are supported per (sel_fn × inference)
# For infer_boot: list supported combinations of inference_target, debias, est_data.
# (Supported nonselection values are looked up directly from `results` inside
# make_diagram(), so no equivalent summary table is needed for those methods.)
supported_boot_vals <- results %>%
  filter(inference == "infer_boot", status %in% c("supported","supported_with_warning")) %>%
  separate_wider_delim(
    boot_label,
    delim = " | ",
    names = c("target", "debias", "est")
  ) %>%
  mutate(
    inference_target = sub("target=", "", target),
    debias = sub("debias=", "", debias),
    estimation_data  = sub("est=", "", est)
  ) %>%
  group_by(sel_fn) %>%
  summarize(
    targets  = list(sort(unique(inference_target))),
    debiases = list(sort(unique(as.character(debias)))),
    est_data = list(sort(unique(estimation_data))),
    .groups  = "drop"
  )

# ── DOT helpers ───────────────────────────────────────────────────────────────
safe_id <- function(x) gsub("[^a-zA-Z0-9_]", "_", x)

node_def <- function(id, label, fill, fontcolor="white",
                     fontsize=10, width=1.9, height=0.45, penwidth=0,
                     border_col=fill) {
  sprintf(
    '  "%s" [label="%s", shape="rectangle", style="filled,rounded",
             fillcolor="%s", fontcolor="%s", fontsize=%d,
             width=%.2f, height=%.2f, penwidth=%.1f, color="%s"]',
    safe_id(id), gsub('"', "'", label), fill, fontcolor,
    fontsize, width, height, penwidth, border_col
  )
}

edge_def <- function(from, to, col, dashed=FALSE, penwidth=1.5) {
  sprintf('  "%s" -> "%s" [color="%s", style="%s", penwidth=%.1f, arrowsize=0.6]',
          safe_id(from), safe_id(to), col,
          if (dashed) "dashed" else "solid", penwidth)
}

# ── make_diagram ───────────────────────────────────────────────────────────────
# Selector flavor nodes are colored by their selector family (sel_palette).
# Inference method nodes and edges use inf_palette — one color per method.
# Terminal option nodes use a faded version of their inference method color.

make_diagram <- function(sel_fn_filter, pair_best, inf_palette, sel_palette,
                         results, supported_boot_vals,
                         splines = "polyline") {
  # splines="ortho" crashes Graphviz maze router (maze.c assertion) on dense
  # graphs. "polyline" is stable at any size. Pass splines="curved" for a
  # slightly prettier look on smaller diagrams if desired.

  pb <- pair_best %>% filter(sel_fn == sel_fn_filter)
  if (nrow(pb) == 0) return(invisible(NULL))

  sel_flavors <- unique(pb$sel_label)
  inf_shown    <- unique(pb$inference)
  sel_col      <- sel_palette[[sel_fn_filter]]

  lines <- c(
    'digraph {',
    sprintf('graph [layout="dot", rankdir="LR", splines="%s",', splines),
    '       bgcolor="white", nodesep=0.35, ranksep=0.8]',
    'node [fontname="Helvetica"]',
    'edge [fontname="Helvetica", fontsize=8]',
    '',
    '// Column 1: selector flavor nodes (each its own color)'
  )

  for (s in sel_flavors) {
    disp <- sub(paste0("^", sel_fn_filter, "\\s*\\|\\s*"), "", s)
    if (disp == s) disp <- s
    lines <- c(lines, node_def(s, disp, fill=sel_col, fontcolor="white",
                               width=2.2, penwidth=0))
  }

  # Inference nodes: colored by inf_palette
  lines <- c(lines, '', '// Column 2: inference method nodes (inf_palette color)')
  for (i in inf_shown) {
    lines <- c(lines, node_def(i, i, fill=inf_palette[[i]], fontcolor="white",
                               width=1.9, penwidth=0))
  }

  # Terminal nodes: one per nonselection value OR one per boot option dimension.
  # These are shared across flavors — the color story is told by the edges.
  lines <- c(lines, '', '// Column 3: terminal option value nodes (neutral)')

  for (i in inf_shown) {
    # Terminal nodes use a faded version of the inference method's own color
    term_fill <- paste0(inf_palette[[i]], "BB")

    if (i == "infer_boot") {
      bv <- supported_boot_vals %>% filter(sel_fn == sel_fn_filter)
      if (nrow(bv) == 0) next
      dims <- list(
        list(id    = paste0(safe_id(i), "_target"),
             label = paste0("inference_target\n",
                            paste(bv$targets[[1]], collapse=" | "))),
        list(id    = paste0(safe_id(i), "_debias"),
             label = paste0("debias\n",
                            paste(bv$debiases[[1]], collapse=" | "))),
        list(id    = paste0(safe_id(i), "_estdata"),
             label = paste0("estimation_data\n",
                            paste(bv$est_data[[1]], collapse=" | ")))
      )
      for (d in dims) {
        lines <- c(lines, sprintf(
          '  "%s" [label="%s", shape="rectangle", style="filled,rounded",
                   fillcolor="%s", fontcolor="white", fontsize=8,
                   width=2.1, height=0.55, penwidth=0]',
          d$id, d$label, term_fill))
      }

    } else if (i == "infer_pipe") {
      # infer_pipe has no option arguments — single terminal node
      lines <- c(lines, sprintf(
        '  "%s" [label="(no options)", shape="rectangle", style="filled,rounded",
                 fillcolor="%s", fontcolor="white", fontsize=8,
                 width=1.6, height=0.42, penwidth=0]',
        paste0(safe_id(i), "_noopts"), term_fill))

    } else {
      # nonselection methods (infer_selective, infer_upsi)
      supported_rows <- results %>%
        filter(sel_fn==sel_fn_filter, inference==i,
               status %in% c("supported","supported_with_warning"))
      nonsel_vals_used <- sort(unique(supported_rows$nonselection))
      if (length(nonsel_vals_used) == 0) next
      for (v in nonsel_vals_used) {
        nid <- paste0(safe_id(i), "_", safe_id(v))
        lines <- c(lines, sprintf(
          '  "%s" [label="%s", shape="rectangle", style="filled,rounded",
                   fillcolor="%s", fontcolor="white", fontsize=8,
                   width=1.9, height=0.42, penwidth=0]',
          nid, v, term_fill))
      }
    }
  }

  # ── Edges: flavor → inference (colored by inference method) ────────────────
  lines <- c(lines, '', '// Flavor → inference edges (inf_palette color)')
  for (s in sel_flavors) {
    infs_ok <- pb %>% filter(sel_label == s) %>% pull(inference) %>% unique()
    for (i in infs_ok) {
      best <- pb %>% filter(sel_label==s, inference==i) %>% pull(best_status)
      dashed <- best == "supported_with_warning"
      lines <- c(lines, edge_def(s, i, inf_palette[[i]], dashed, penwidth=1.6))
    }
  }

  # ── Edges: inference → terminal (colored by inference method)
  # Drawn once per inference method (not per flavor) since color is per-method.
  lines <- c(lines, '', '// Inference → terminal edges (inf_palette color)')
  for (i in inf_shown) {
    col <- inf_palette[[i]]

    if (i == "infer_boot") {
      bv <- supported_boot_vals %>% filter(sel_fn == sel_fn_filter)
      if (nrow(bv) == 0) next
      for (sfx in c("_target","_debias","_estdata")) {
        lines <- c(lines, edge_def(i, paste0(safe_id(i), sfx), col, penwidth=1.0))
      }

    } else if (i == "infer_pipe") {
      lines <- c(lines, edge_def(i, paste0(safe_id(i), "_noopts"), col, penwidth=1.0))

    } else {
      # nonselection methods: one edge per supported nonselection value
      supported_rows <- results %>%
        filter(sel_fn==sel_fn_filter, inference==i,
               status %in% c("supported","supported_with_warning"))
      for (v in sort(unique(supported_rows$nonselection))) {
        lines <- c(lines, edge_def(i, paste0(safe_id(i),"_",safe_id(v)), col, penwidth=1.0))
      }
    }
  }

  lines <- c(lines, '}')
  grViz(paste(lines, collapse="\n"))
}
```



### `select_stepwise_ic`  (penalty × direction)

```{r diag-stepwise, fig.width=11, fig.height=6}
make_diagram("select_stepwise_ic", pair_best, inf_palette, sel_palette,
             results, supported_boot_vals)
```


**Interpretation:** All six AIC/BIC × direction combinations support
`infer_upsi` and `infer_boot`.  `infer_selective` is only available for
`direction = "forward"`: the conditioning it performs is defined in terms of a
forward-stepwise path, and `selectiveInference::fsInf()` implements that case
only.  `infer_pipe` is not available for stepwise selection.

### `select_glmnet`  (alpha × lambda: ridge/elastic net/lasso × best/compact)

```{r diag-glmnet, fig.width=11, fig.height=5}
make_diagram("select_glmnet", pair_best, inf_palette, sel_palette,
             results, supported_boot_vals)
```


**Interpretation:** `infer_selective` (shown in blue) is only connected for
`alpha = 1.0` (lasso) flavors.  Ridge (`alpha = 0`) and elastic net
(`alpha = 0.5`) can only be paired with `infer_upsi` or `infer_boot`.
All `lambda` options work equivalently for supported inference methods.


### `select_ncvreg`  (penalty × lambda: MCP/SCAD/lasso × best/compact)

```{r diag-ncvreg, fig.width=11, fig.height=5}
make_diagram("select_ncvreg", pair_best, inf_palette, sel_palette,
             results, supported_boot_vals)
```

**Interpretation:** All three penalties (MCP, SCAD, lasso) and both `lambda`
options support `infer_upsi`, `infer_boot`, and `infer_pipe`.  `infer_selective`
is not available for `ncvreg` selectors — note that this applies even to
`penalty = "lasso"`, because selective inference needs the `glmnet`
parameterization of the lasso fit rather than `ncvreg`'s.


# All probed combinations

The table below is the raw output behind the matrix and diagrams: one row per
selector flavor × inference method × option combination, with the error or
warning message where one was produced.  Use the filter boxes to narrow it
down.

```{r full-table}
datatable(
  results %>%
    rename(`selector fn`=sel_fn, `selector params`=sel_params,
           `selector label`=sel_label, `boot options`=boot_label),
  rownames=FALSE, filter="top",
  options=list(pageLength=15, scrollX=TRUE),
  caption="All probed combinations — filter by any column"
)
```

# For developers

This vignette is designed to update itself automatically as new methods are
added to the package.  The probing logic in the `probe` chunk calls each
function via `get(fn_name, envir=getNamespace("selectInferToolkit"))`, so any
newly exported `select_*` or `infer_*` function is discovered without editing
this file.  Tuning *options*, however, are not discoverable, so a new method
with its own options needs one of the following small edits.

To add a **new selection method** with tuning parameters, add one entry to
`sel_grids` in the `candidates` chunk:

```r
select_mymethod = expand.grid(
  option1 = c("a", "b"),
  option2 = c(TRUE, FALSE),
  stringsAsFactors = FALSE
)
```

To add a **new inference method that uses `nonselection=`**, add its name to
`inf_fns_nonsel` in this vignette's Rmd.

To add a **new inference method with its own option grid** (like `infer_boot`), add a dedicated probe block in the `probe` chunk following the pattern used for `grid_boot`.

The compatibility matrix and flow diagrams will include the new method
automatically on the next render.






